---
title: "Project 2 - Karen Lopez"
date: "October 16, 2020"
output:
  rmarkdown::github_document:
    toc: true
knit: (function(input, ...){
  rmarkdown::render(
    input,
    output_file="README.md",
    envir=globalenv()
  )})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction  
This project uses the bike sharing data set, day.csv, that's located *[here](https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset)* and contains 731 observations with 15 attributes. For modeling, the response variable is the count of total rental bikes rented (*cnt*) and 12 of the 14 remaining variables will be considered for predictor variables (*casual* and *registered* are omitted). The 12 remaining variables include date, season, year, month, holiday, weekday, working day, weather, temperature, feeling temperature, humidity, and wind speed.  
Purpose of analysis:  
Methods I'll use:  
To begin, necessary libraries are loaded so their functions are accessible and global variables are set.  
```{r libraries, message=FALSE}
library(readr)
library(caret)
library(knitr)
library(corrplot)
library(dplyr)
num <- 12
```

## Data  
The dataset is read in using a relative path and saved as an object. The data is then randomly split into a training and testing set, where 70% of the data goes into the training set and the remaining 30% goes into the testing set.  
```{r datasets, warning=FALSE}
bikeData <- read_csv("Bike-Sharing-Dataset/day.csv") #read in data
bikeData$cnt <- as.factor(bikeData$cnt) #convert response to factor

#create partitions in data indexes with 70% going in the training set
set.seed(num)
train <- sample(1:nrow(bikeData), size = nrow(bikeData)*0.7)
test <- dplyr::setdiff(1:nrow(bikeData), train)

#create train and test data set with split indexes
bikeDataTrain <- bikeData[train, ]
bikeDataTest <- bikeData[test, ]

#kable(bikeDataTrain) #check for 70% of data -> 511
#kable(bikeDataTest) #check for 30% of data -> 220

kable(head(bikeDataTrain)) #preview train data
kable(head(bikeDataTest)) #preview test data
```

## Summarizations  
summary statistics and plot with some general explanations  
```{r summaries}

```

## Modeling  
fit some models (two)  
```{r models}

```

## Automation & Blog Post  
automate so it creates report for each day of the week  
```{r automation}

```

